---
title: "Data Wrangling On Contraception Method Dataset"
author: "Laura Cortes"
date: "September 15, 2017"
output: html_document
---  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Contraception Method - Getting Data Ready for Analysis

```{r echo = FALSE}

library(ggplot2)
library(tidyverse)
library(dplyr)

#Read in the data

cmc <- read.table("https://archive.ics.uci.edu/ml/machine-learning-databases/cmc/cmc.data", sep=",")

# Attribute Information:
# 
#    1. Wife's age                     (numerical)
#    2. Wife's education               (categorical)      1=low, 2, 3, 4=high
#    3. Husband's education            (categorical)      1=low, 2, 3, 4=high
#    4. Number of children ever born   (numerical)
#    5. Wife's religion                (binary)           0=Non-Islam, 1=Islam
#    6. Wife's now working?            (binary)           0=Yes, 1=No
#    7. Husband's occupation           (categorical)      1, 2, 3, 4
#    8. Standard-of-living index       (categorical)      1=low, 2, 3, 4=high
#    9. Media exposure                 (binary)           0=Good, 1=Not good
#    10. Contraceptive method used     (class attribute)  1=No-use 
#                                                         2=Long-term
#                                                         3=Short-term

#Add column names
colnames(cmc) <- c("WifeAge","WifeEdu","HusbandEdu","Children","WifeReligion",
                   "WifeWorks","HusbandOcc","SOLIndex","MediaExp","CMethod")

```
## Reviewing Data  
Below we will take a look at the dataset, identifying columns' names and types. Also we will look for missing and extreme values. In the case we find any, we will replace them with the column's mean value.   

```{r echo = TRUE}

#Take a look at the data structure
str(cmc)
#Only 2 of the 10 variables are numeric, the rest are binary/categorical

#Convert to factor non-numeric variables 
cmc[2:3] <- lapply(cmc[2:3], factor)
cmc[5:10] <- lapply(cmc[5:10], factor)

#Looking for missing values
sum(is.na(cmc))

#Looking for extreme values and outliers
#Take a look at the statistics
summary(cmc)
#Take a look at the boxplot to identify outliers
boxplot(cmc)

#Review in detail columns with apparent extreme values / outliers
#Check husband's occupation's values
summary(cmc$HusbandOcc)
#Check standard of living index's values
summary(cmc$SOLIndex)
#Check number of children's values
hist(cmc$Children)
boxplot(cmc$Children)
plot(cmc$WifeAge, cmc$Children)

#Children column seems to have extreme values
#Check how many rows has number of children greater than 10
sum(cmc$Children > 10)
#Find indice of extreme values / outliers
ind <- which(cmc$Children > 10)
#Look at the full rows
cmc[ind, ]

#Set Children to an appropriate value, in our case, we will use the mean value
# Calculate the mean of the Children column and round it
children_mean <- round(mean(cmc$Children, na.rm = TRUE, trim = 0))

# Set Children with extreme values to mean
cmc$Children[ind] <- children_mean
cmc[ind, ]

#Take a look at Children column after cleaning up extreme values
hist(cmc$Children)
boxplot(cmc$Children)
plot(cmc$WifeAge, cmc$Children) 

#The dataset is now clean 

#Create CSV file
write.csv(cmc, file="cmc_clean.csv", row.names = FALSE)

```
#### Conclusion  
+ The dataset didn't present missing values
+ Only one of the columns seemed to have extreme values: Children 
+ Children's extreme values (> 10), were replaced by the mean value
+ Columns' names were changed to make them more readable
+ A CSV file named "cmc_clean.csv" was created with the clean dataset



 

