---
title: "Exploring_CM"
author: "Laura Cortes"
date: "September 14, 2017"
output: html_document
--- 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Contraception Method - Exploratory Analysis

```{r echo = FALSE}

# Contraception Method Dataset

library(ggplot2)
library(tidyverse)

#Read in the data

cmc <- read.table("https://archive.ics.uci.edu/ml/machine-learning-databases/cmc/cmc.data", sep=",")

# Attribute Information:
# 
#    1. Wife's age                     (numerical)
#    2. Wife's education               (categorical)      1=low, 2, 3, 4=high
#    3. Husband's education            (categorical)      1=low, 2, 3, 4=high
#    4. Number of children ever born   (numerical)
#    5. Wife's religion                (binary)           0=Non-Islam, 1=Islam
#    6. Wife's now working?            (binary)           0=Yes, 1=No
#    7. Husband's occupation           (categorical)      1, 2, 3, 4
#    8. Standard-of-living index       (categorical)      1=low, 2, 3, 4=high
#    9. Media exposure                 (binary)           0=Good, 1=Not good
#    10. Contraceptive method used     (class attribute)  1=No-use 
#                                                         2=Long-term
#                                                         3=Short-term

#Add column names
colnames(cmc) <- c("WifeAge","WifeEdu","HusbandEdu","Children","WifeReligion",
                   "Working","HusbandOcc","SOLIndex","MediaExp","Method")

#Take a look at the data structure
str(cmc)

#Convert to factor variables (Only 2 out of the 10 variables are numerical)
cmc[2:3] <- lapply(cmc[2:3], factor)
cmc[5:10] <- lapply(cmc[5:10], factor)

#Exploratory Data Analysis
#Vertical facets

```
## Density Curve for Number of Children
```{r echo = TRUE}
ggplot(cmc, aes(x=Children)) + geom_histogram(aes(y=..density..),binwidth = 3)+
  stat_function(fun=dnorm, colour='red',
                args=list(mean=mean(cmc$Children, na.rm=T),
                          sd=sd(cmc$Children, na.rm=T)))


```


## Wife's Age and Number of Children
```{r echo = TRUE}
ggplot(cmc, aes(x = WifeAge, y = Children, color = Method)) + 
  geom_jitter(alpha = 0.9) + facet_grid(. ~ Method)

``` 
  
  
### Comments:
On the previous plots we observe there are women with more than 10 children.

On the second plot we observe that the women with more than 10 children are 35 years old, or older, which make some sense (?).  

Should we consider the children over 12 as outliers (4 cases)? 



 

 

